\section*{Acknowledgements}

I would like to thank professor Quoilin first, for his mentoring during this thesis and his advices. Then Jade, for her collaboration on the linkings between the created model and MEDEAS.

Computational resources have been provided by the Consortium des Équipements de Calcul Intensif (CÉCI), funded by the Fonds de la Recherche Scientifique de Belgique (F.R.S.-FNRS) under Grant No. 2.5020.11 and by the Walloon Region. I am also grateful to the cluster administrator, M. Collignon, for his patience and valuable advices.

The use of the popular tool ChatGPT \cite{chatgpt} is also to be mentioned, although primarily taken advantage of for proof-reading and suggestions.

Thanks should also go to my mother, and everyone to which I talked about my thesis, what sometimes led me to get a new perspective on the work as I had to explain it in a new, concise and understandable way.

\newpage

\section*{Abstract}

This master thesis explores the application of a surrogate model to achieve the integration of electricity network constraints into integrated assessment models, that attempt to predict the evolution of the main features of the society, including economic, demographic, climatic and energy factors.

Tools were developped to run Dispa-SET, a dispatch model for the European electricity network, to build a dataset of simulations on adequate sample points. This dataset is then used to train the surrogate model employing appropriate machine learning methods. Finally, this model is integrated into the MEDEAS IAM, connecting their variables in a meaningful way, and the resulting model is run for different scenarios.

Results show that integrating the model in MEDEAS leads to a lower prediction of the production of variable renewable energy sources compared to the initial MEDEAS output.

\vspace{1cm}
Keywords: surrogate model, Dispa-SET, MEDEAS, machine learning, VRES

\newpage

\tableofcontents

\newpage

\section*{Keywords}

\addcontentsline{toc}{section}{Keywords}

\begin{table}[h!]
    \begin{tabular}{p{3cm} p{9cm}}
        AF & Availability factor \\
        ANN & Artificial neural network \\
        BAU & Business as ususal \\
        CF & Capacity factor \\
        CÉCI & Consortium des Équipements de Calcul Intensif \\
        DLL & Dynamically loaded library \\
        ENS & Energy not served \\
        EROI & Energy return on investment \\
        EU & European union \\
        GAMS & General algebraic modeling language \\
        IAM & Integrated assessment model \\
        LP & Linear programming \\
        LHS & Latin hypercube sampling \\
        MILP & Mixed integer linear programming \\
        MLT & Mid-level transition \\
        MLP & Multi-layer perceptron \\
        MTS & Mid-term scheduling \\
        NTC & Net transfer capacity \\
        OLT & Optimal level transition \\
        P2H & Power to heat \\
        PHS & Pumped hydro-storage \\
        PV & Photovoltaic \\
        RES & Renewable energy sources \\
        SLURM & Simple utility for Resource Management \\
        VRES & Variable renewable energy sources \\
    \end{tabular}
\end{table}

\newpage
